Quantifying the importance of a tree prior in Bayesian phylogenetics

### 2019-04-10 article.tex
Reviewed description

\richel{
  I think this is a usefull paragraph, adding some depth to this paper,
  but it could use some rewording
}
There have been multiple attemps to investigate the importance of tree
prior choice. For example, recently Sarver et al., [\cite{sarver2019choice}] 
show that the choice of tree prior does not 
substantially affect phylogenetic inferences of diversification rates.
Also recently, Duchene et al. [\cite{duchene2018phylodynamic}] released
a BEAST2 package to assess how well posterior predictive simulations
recover a given tree when using the standard diversification models.
These investigations show how current diversification models compare
to one another, but these do not help assess the importance of a 
new tree prio.

@@

Using a twin tree serves as a control: both $\tau_{\mathit{P}}$ and 
$\tau_{\mathit{I}}$ will produce an error distribution. 
These errors are unavoidably accumulated in the \verb;pirouette; pipeline. 
All the factors that can cause such errors are: the branching 
times $\vec{t}$ of the tree (as a consequence of the tree prior), its topology 
$\psi$, the stochasticity induced by simulating the alignment (step 1a 
$\rightarrow$ 2a in Fig.~\ref{fig:pipeline}) and the stochasticity induced by 
the Bayesian inference (step 2a $\rightarrow$ 4a in Fig.~\ref{fig:pipeline}). 
By imposing the same topology and replicating the pipelines a sufficient number 
of times we can sort out all the possible causes that are not a direct 
consequence of the choice of the tree prior.

Potential differences in shape of error distributions will therefore be 
consequences of different tree priors for the two parallel pipelines: the use 
of prior $\mathit{P}$ for the original pipeline and of prior $\mathit{I}$ for 
the twin pipeline. In the second case, though, such error distribution is not 
influenced by the mismatch between how the tree is generated and how it is 
analyzed by BEAST2.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Installation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
  \centering
  \begin{tabular}{ | l | l | }
    \hline
    \textbf{Name} & \textbf{Description} \\
    \hline
    \verb;pir_run; & Run \verb;pirouette; \\
    \verb;create_pir_params; & Create the \verb;pirouette; parameters \\
    \hline
    \verb;create_alignment_params; & Create the alignment parameters \\
    \verb;create_twinning_params; & Create the twinning parameters \\
    \verb;create_experiment; & Create one experiment \\
    \verb;create_error_measure_params; & Create the error measurement parameters \\
    \hline
  \end{tabular}
  \caption{pirouette's main functions.}
  \label{tab:functions}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The error distribution returned by \verb;pir_run; is stored as a data frame. 
\richel{
  I suggest \texttt{pir\_run} to return a list of run experiments,
  as described in Issue 225, 
  \url{https://github.com/richelbilderbeek/pirouette/issues/225}
}
It can be directly plotted using \verb;pir_plot;:

>>>>

There have been multiple attempts to investigate the importance of tree
prior choice. For example, recently Sarver et al., [\cite{sarver2019choice}] 
show that the choice of tree prior does not 
substantially affect phylogenetic inferences of diversification rates.
Also recently, Duchene et al. [\cite{duchene2018phylodynamic}] released
a BEAST2 package to assess how well posterior predictive simulations
recover a given tree when using the standard diversification models.
These investigations show how current diversification models compare
to one another, but these do not help assess the importance of a 
new tree prio.

@@

Using a twin tree serves as a control: both $\tau_{\mathit{P}}$ and 
$\tau_{\mathit{I}}$ will produce an error distribution. 
These errors unavoidably accumulate in the \verb;pirouette; pipeline,
in both the true and twin tree: either tree may be an unlikely representative
of the underlying tree prior, the alignment simulation may give different
numbers of mutations for a same evolutionary time, 
and the MCMC will sample different states.
These errors can be equalized by imposing a same tree topology and by
replicating the pipelines a sufficient number.

The left-over differences in error distributions will be 
consequences of different phylogeny and tree priors for the two parallel 
pipelines: the main pipeline uses tree $\tau_{\mathit{P}}$ and
possibly mismatched prior $\mathit{P}$, where the twin pipeline
uses tree $\tau_{\mathit{I}}$ for a matching prior $\mathit{I}$.
So, for the twin route, the resulting error distribution cannot be 
attributed to a mismatch between true tree prior and tree prior used in
inference.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Installation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
  \centering
  \begin{tabular}{ | l | l | }
    \hline
    \textbf{Name} & \textbf{Description} \\
    \hline
    \verb;pir_run; & Run \verb;pirouette; \\
    \verb;pir_plot; & Show the \verb;pirouette; results as a plot \\
    \verb;create_pir_params; & Create the \verb;pirouette; parameters \\
    \hline
    \verb;create_alignment_params; & Create the alignment parameters \\
    \verb;create_twinning_params; & Create the twinning parameters \\
    \verb;create_experiment; & Create one experiment \\
    \verb;create_error_measure_params; & Create the error measurement parameters \\
    \hline
  \end{tabular}
  \caption{pirouette's main functions.}
  \label{tab:functions}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The error distribution returned by \verb;pir_run; is stored as a data frame. 
It can be directly plotted using \verb;pir_plot;:
